---
title: "Lab 7 exercise"
output: word_document
---

```{r,include=FALSE}
options(digits=3)
```

```{r setup, message=FALSE,warning=FALSE,results='hide'}
library(mosaic)
library(tidyverse)
```

Read in "NHANES" dataset and take the adult (age $18+$ subset: (change the code below so that it points to where you have nhanes.csv saved)

```{r, results='hide'}
nhanes <- read.csv('nhanes.csv' , header = T)
### only adults
nhanes <- filter(nhanes,age>=18)
```

# Exercise:

Now study the difference of the mean weight between adults with prehypertension and adults with SPB Stage 1 Hypertension.

Here are datasets of people with PreHypertension and Stage 1 Hypertension, respectively:

```{r}
PreHyp <- filter(nhanes,sbpClass=='Prehypertension')

stage1 <- filter(nhanes,sbpClass=='Stage 1 Hypertension')

w1 <- stage1$weight; #Group1: weights for normal SBP;
w2 <- PreHyp$weight; #Group2: weights for prehypertension;
```

-   Formulate a null and alternative hypothesis for a two-sided hypothesis test for the difference in mean weight between adults who are Prehypertensive and those with Stage 1 hypertension (you may use words instead of math, if you prefer):

$H_0:$ The population mean weight of people with sbpClass= Normal is equal to the population mean weight of people with sbpClass= Prehypertension

$H_a:$ The population mean weight of people with sbpClass= Normal is not equal to the population mean weight of people with sbpClass= Prehypertension

We will test $H_0$ at level $\alpha=0.05$.

-   What is the difference in sample means between the two populations?

```{r}
y1 <- mean(w1); #sample mean of group1. 
y2 <- mean(w2); #sample mean of group2.
y1
y2
```

-   What is the standard error of the difference in sample means?

```{r}
s1 <- sd(w1); #sample SD
s2 <- sd(w2); 

n1 <- length(w1); #sample size
n2 <- length(w2); #sample size

V1 <- s1^2/n1; #sampling variance of sample mean of group 1.
V2 <- s2^2/n2; #sampling variance of sample mean of group 2.

Vdiff <- V1+V2 # sampling variance of difference of two sample means

SE <- sqrt(Vdiff); #Standard error for the difference of sample means.
SE
```

-   What is the T-statistic (ratio test statistic) to test $H_0$ versus $H_a$?

```{r}
## put your answer here
T=(y1-y2)/SE
T

```

-   Use the same test statistic to estimate a p-value testing $H_0$

```{r}
## put your answer here
2*pt(-abs(T),df=min(n1,n2)-1)
```

-   What is the correct multiplier (i.e. critical value) for a 95% confidence interval for $\mu_{preH}-\mu_{Stg1}$?

```{r}
## put your answer here
alpha <- 1-0.95
mult <- qt(1-alpha/2,df=min(n1,n2)-1)
mult
```

-   Use the difference in sample means, standard error, and critical value you calculated to estimate a 95% confidence interval for $\mu_{preH}-\mu_{Stg1}$

```{r}
## put your answer here
mean(w1)-mean(w2)+c(-1,1)*mult*SE
```

-   Use function t.test to estimate a p-value testing $H_0$ and a 95% confidence interval Does it (roughly) agree?

```{r}
#put your answer here
t.test(w1, w2,conf.level = 0.95); #Check the result "95 percent
```

-   What is your decision?

we will reject the null hypothesis because p\< alpha

-   What can you conclude about the mean weight of people with pre-hypertension verus stage-1 hypertension?

There is no evidence that the mean weight of people with pre-hypertension us equal to the mean weigth of people with stage-1 hypertension
